---
title: "MPG difference between automatic and manual transmission"
author: "X. Zeng"
date: "August 19, 2015"
output: pdf_document
---

## Introduction
Looking at a data set of a collection of cars, people in Motor Trend, a magazine about the automobile industry, are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”

"Quantify the MPG difference between automatic and manual transmissions"

In this report, I'll use regression models to answer the questions.

## Exploratory data analysis

First, let's load the data:

```{r, echo=TRUE}
library(ggplot2)
data(mtcars)
head(mtcars)
```

Then, let's use a boxplot to see the mpg difference between automatic (`am=0`) and manual (`am=1`) transmission. For the figure (in Appendix), we can see there are some difference between to two mpgs. 

## Regression 

Now, let's use two linear regression models to see if there's some difference between different transimission systems. 

```{r, echo=FALSE}
df1 <- mtcars[mtcars$am==1,]
df2 <- mtcars[mtcars$am==0,]
dim(df1)
dim(df2)
ind <- sample(1:nrow(df2), nrow(df1))
df2 <- df2[ind,]
fit1 <- lm(mpg~.-am, data=df1)
fit2 <- lm(mpg~.-am, data=df2)
summary(fit1)
summary(fit2)
```

As we can see from the summary, the model generally represent the variation of mpg for both cases. Now, let's use t-test to see if there's any difference between the two fitted models.

```{r, echo=TRUE}
t.test(fit1$fitted.values,fit2$fitted.values)
```

Apparently, the p-value is much less than 0.05. So we can reject the null hyposis that there's no difference between the two means related to automatic and manual transmission.

## Conclusion

The MPG difference between manual and automatic transmissions is `r mean(fit1$fitted.values-fit2$fitted.values)`. That is, manual has higher mpg than automatic transmission.

## Appendix

```{r, echo=FALSE}
ggplot(mtcars, aes(as.factor(am),mpg))+geom_boxplot()
```

```{r, echo=FALSE}
plot(fit1$mpg, fit1$residuals, xlab="mpg",ylab="residuals", main="residuals for automatic")
```

```{r, echo=FALSE}
plot(fit2$mpg, fit2$residuals, xlab="mpg",ylab="residuals", main="residuals for manual")
```

